---
title: "Project"
author: Jonas Noury
format: html
execute:
  echo: false
---
I plan to use this project to see correlation between inflation and transfer market prices over the year(soccer). I will be taking data from transfermarkt.com and inflation data from the World Bank. I will be using R to analyze the data and create visualizations to show any trends or correlations that may exist.
# Project

```{r}
#| label: setup
#| message: false

library(tidyverse)
```

## Transfer Market Values Over Time

```{r}
#| label: plot-transfers
#| message: false

# Read transfer data
transfers <- read_csv("data/transfers.csv")

# Calculate average market value per year and plot starting from 2005
transfers |>
  mutate(year = year(transfer_date)) |>
  filter(!is.na(market_value_in_eur), market_value_in_eur > 0, year >= 2005, year <= 2024) |>
  group_by(year) |>
  summarize(mean_market_value = mean(market_value_in_eur)) |>
  ggplot(aes(x = year, y = mean_market_value / 1e6)) +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(color = "steelblue", size = 2) +
  scale_x_continuous(breaks = seq(2005, 2024, by = 3)) +
  scale_y_continuous(labels = scales::label_dollar(prefix = "€", suffix = "M")) +
  labs(
    title = "Average Player Market Value Over Time",
    subtitle = "Data from Transfermarkt (2005–2024)",
    x = "Year",
    y = "Average Market Value (Millions €)"
  ) +
  theme_minimal()
```

## Consumer Price Index Over Time

```{r}
#| label: plot-cpi
#| message: false

# Read World Bank CPI data (skip 4 metadata lines)
cpi <- read_csv("data/world_bank_cpi.csv", skip = 4)

# Reshape year columns to long format and plot global median CPI from 2005
cpi |>
  pivot_longer(
    cols = matches("^[0-9]{4}$"),
    names_to = "year",
    values_to = "cpi_value"
  ) |>
  mutate(year = as.numeric(year)) |>
  filter(year >= 2005, year <= 2024, !is.na(cpi_value)) |>
  group_by(year) |>
  summarize(median_cpi = median(cpi_value)) |>
  ggplot(aes(x = year, y = median_cpi)) +
  geom_line(color = "firebrick", linewidth = 1) +
  geom_point(color = "firebrick", size = 2) +
  scale_x_continuous(breaks = seq(2005, 2024, by = 3)) +
  labs(
    title = "Global Consumer Price Index Over Time",
    subtitle = "Median across countries from World Bank (2010 = 100)",
    x = "Year",
    y = "Consumer Price Index (2010 = 100)"
  ) +
  theme_minimal()
```

## Consumer Price Index Observations Since 2005

```{r}
#| label: plot-cpi-observations
#| message: false

# Plot every available country-year CPI observation since 2005
cpi |>
  pivot_longer(
    cols = matches("^[0-9]{4}$"),
    names_to = "year",
    values_to = "cpi_value"
  ) |>
  mutate(year = as.numeric(year)) |>
  filter(year >= 2005, year <= 2024, !is.na(cpi_value), cpi_value > 0) |>
  ggplot(aes(x = year, y = cpi_value)) +
  geom_jitter(
    width = 0.15,
    height = 0,
    alpha = 0.22,
    size = 1.2,
    color = "#2878b5"
  ) +
  stat_summary(
    fun = median,
    geom = "line",
    aes(group = 1),
    color = "#d1495b",
    linewidth = 1.1
  ) +
  scale_x_continuous(breaks = seq(2005, 2024, by = 3)) +
  scale_y_log10(labels = scales::label_number()) +
  labs(
    title = "Country-Level Consumer Price Index Observations",
    subtitle = "Thousands of World Bank country-year observations since 2005; red line shows the yearly median",
    x = "Year",
    y = "Consumer Price Index (2010 = 100, log scale)"
  ) +
  theme_minimal()
```
